---
title: "Australian Census 2011 - Cycling to work"
author: "coolbutuseless"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Vignette Title}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r message=FALSE}
library(knitr)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(rgdal) # port install gdal, R CMD INSTALL rgdal_0.9-1.tar.gz
library(pryr)
library(rgeos) # gSimplify  - port install geos; R CMD INSTALL rgeos_0.3-8.tar.gz
library(ggmap)
library(leaflet)

library(AuCensus2011)
```


```{r}
#-----------------------------------------------------------------------------
# The extents of the captial cities was chosen arbitrarily
#-----------------------------------------------------------------------------
#                 TOP             LEFT          BOTTOM        RIGHT
australia <- list(max_lat=-10    , min_long=112  , min_lat=-45  , max_long=155)
brisbane  <- list(max_lat=-27.4  , min_long=152.9, min_lat=-27.7, max_long=153.2)
sydney    <- list(max_lat=-33.75 , min_long=151.1, min_lat=-34.0, max_long=151.3)
melbourne <- list(max_lat=-37.7  , min_long=144.7, min_lat=-37.9, max_long=145.1)
adelaide  <- list(max_lat=-34.76 , min_long=138.45,min_lat=-35.0, max_long=138.75)
perth     <- list(max_lat=-31.7  , min_long=115.7 ,min_lat=-32.5, max_long=116.13)
canberra  <- list(max_lat=-35.19 , min_long=149.0 ,min_lat=-35.34,max_long=149.2)
hobart    <- list(max_lat=-42.82 , min_long=147.26,min_lat=-42.9 ,max_long=147.38)
darwin    <- list(max_lat=-12.32 , min_long=130.80,min_lat=-12.48,max_long=130.91)
```


Choose a statistical level and load the data
------------------------------------------------------------------------------

```{r}
this_level <- 'SSC'
b46 <- AuCensus2011::read_abs("BCP", "B46", this_level)
knitr::kable(b46[1:10, 1:5], caption="Table B46 Method of Travel to Work (raw census data)")
```

Filter data and convert to long format with augmented columns
```{r}
b46 %<>% abs_wide_to_long %>%
    filter(gender == 'Persons') %>%
    select(SSC_CODE, method, method_count, count)
knitr::kable(head(b46), caption="Table B46 Method of Travel to Work. Filtered, long data with augmented columns")
```


Summarise the cycling data and work out the rate (% working population who travelled to work)
------------------------------------------------------------------------------
```{r}
cycled <- b46 %>% filter(method=='Bicycle') %>% select(1, Bicycle=count)
total  <- b46 %>% filter(method_count %in% c('Total', 'Did_not_go_to_work', 'Worked_at_home')) %>%
    spread(method_count, count)

cycled %<>% inner_join(total) %>% mutate(cycled = 100 * Bicycle / (Total - Did_not_go_to_work - Worked_at_home))

knitr::kable(head(cycled), caption="Calculated cycling rate by suburb")
```


Load the ASGS statistical boundaries
------------------------------------------------------------------------------
```{r}
boundaries <- AuCensus2011::read_asgs_shapefile(level=this_level)
# boundaries <- asgs_simplify_boundaries(boundaries)
knitr::kable(head(boundaries@data), caption="statistical oundaries metadata")
```


ASGS allocation (statistical area hierarchy)
------------------------------------------------------------------------------
The `asgs.sa1.alloc` data.frame contains the allocations (correspondences) between the different statistical areas.

This can be used to filter and select regions. e.g.

* All suburbs in the GCCSA of 'Great Brisbane'
* All 'SA2' areas within a particular 'SA3' area.
* Find which suburbs are covered by a particular postcode.

TODO: `asgs.sa1.alloc` and `asgs.mb.alloc` should probably be another vigneete.

```{r}
knitr::kable(head(asgs.sa1.alloc[1:10, 1:10]), caption="asgs.sa1.alloc allocation (correspondence) table")
```


Plot a subset of all the suburbs in the GCCSA area of 'Greater Melbourne'
------------------------------------------------------------------------------
```{r fig.width=6, fig.height=6}
# Use allocation table to get all suburbs within GCCSA of Melbourne
melbourne_suburbs <- asgs.sa1.alloc %>% filter(GCC_NAME11=='Greater Melbourne') %$% SSC_NAME %>% unique %>% sort
plot(boundaries[boundaries$SSC_NAME %in% melbourne_suburbs,])
```


Trim the boundaries to the bounding box around Melbourne
------------------------------------------------------------------------------
```{r}

if (FALSE) {
    # Subset plotting region by bounding box
    boundaries <- asgs_select_bbox(boundaries, 
                                   melbourne$min_long, melbourne$min_lat,
                                   melbourne$max_long, melbourne$max_lat)
} else {
    # Subset plotting region, but choosing suburbs somehow.
    # Here, use asgs.sa1.alloc to find all the suburbs in Melbourne
    melbourne_suburbs <- asgs.sa1.alloc %>% filter(GCC_NAME11=='Greater Melbourne') %$% SSC_NAME %>% unique %>% sort
    boundaries <- boundaries[boundaries$SSC_NAME %in% melbourne_suburbs,]
}

```


Basic ggplot
------------------------------------------------------------------------------
```{r fig.width=8, fig.height=6}
plot.df <- merge_asgs_abs(boundaries, cycled)
knitr::kable(head(plot.df), caption="plot-ready data set")

if (length(unique(plot.df$group)) > 1) {
    m <- ggplot(plot.df) +
        geom_polygon(aes(x=long, y=lat, group=group, fill=cycled)) +
        coord_fixed()
    print(m)
} else {
    
    m <- ggplot(plot.df) +
        geom_polygon(aes(x=long, y=lat, group=group), fill='#333333') +
        coord_fixed()
    print(m)
}
```

ggmap plotting
------------------------------------------------------------------------------
```{r message=FALSE, fig.width=8, fig.height=6}
# Find the bounding box around what's left of the boundary
region_bbox <- sp::bbox(boundaries)
# ask ggmap to get the map data within this boundary
amap <- ggmap::get_map(location = region_bbox, source = "stamen", maptype = "toner", crop = T)

if (length(unique(plot.df$group)) > 1) {
    p <- ggmap(amap) + geom_polygon(data=plot.df, aes(x=long, y=lat, group=group,  fill=cycled  ), alpha=0.9) + coord_fixed()
    print(p)
} else {
    p <- ggmap(amap) + geom_polygon(data=plot.df, aes(x=long, y=lat, group=group), fill='#333333', alpha=0.9) + coord_fixed()
    print(p)
}

```


ggmap plotting (boundaries only)
------------------------------------------------------------------------------
```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
region_bbox <- sp::bbox(boundaries)
amap <- ggmap::get_map(location = region_bbox, source = "stamen", maptype = "toner", crop = T)

p <- ggmap(amap) + geom_polygon(data=plot.df, aes(x=long, y=lat, group=group), colour='red', fill=NA) + coord_fixed()
print(p)
```


sp plotting
------------------------------------------------------------------------------
```{r fig.width=8, fig.height=6}
# http://r-video-tutorial.blogspot.ch/2015/05/global-economic-maps.html
sp::spplot(boundaries, "SQKM")
```

sp plotting by label
------------------------------------------------------------------------------
```{r fig.width=8, fig.height=6}
library(RColorBrewer)
sp::spplot(boundaries, "CONF_VALUE",
           col.regions = colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(8),
           col = "white"  # colour of boundary lines
)
```

leaflet plot
------------------------------------------------------------------------------

```{r fig.width=8, fig.height=6}
leafletboundaries      <- boundaries
leafletboundaries@data <- boundaries@data %>% left_join(cycled)

# Colour Brewer names:
# Sequential: Blues BuGn BuPu GnBu Greens Greys Oranges OrRd PuBu PuBuGn PuRd Purples RdPu Reds
#             YlGn YlGnBu YlOrBr YlOrRd
# Diverging: BrBG PiYG PRGn PuOr RdBu RdGy RdYlBu RdYlGn Spectral
pal <- colorNumeric("Blues", leafletboundaries@data$cycled)

leaflet(boundaries) %>% 
    addTiles %>% 
    addPolygons(stroke=TRUE,
                color       = 'black',
                weight      = 1.5,
                fillOpacity = 0.8,
                popup       = leafletboundaries@data$SSC_NAME,
                fillColor   = ~pal(leafletboundaries@data$cycled)
    ) %>%
    addLegend(position = "bottomright",
              pal      = pal,
              values   = seq(0, 20, 2.5),
              opacity  = 1,
              title    = "% of commuters who went by bike<br>2011 Census. ABS data")

```


